---
title: "test"
output: pdf_document
date: "2022-10-05"
---

```{r}
pokemon <- read.csv("cleaned_pokemon.csv")
```


```{r}
library(dplyr)
library(ggplot2)

pokemon %>% 
  ggplot(aes(type1,fill = factor(is_legendary))) + 
  geom_bar()
```

```{r}
library(treemap)
 
# Build Dataset
#group <- pokemon$type1
#subgroup <- pokemon$type2
#value <- data.frame(table(pokemon$type1))
#data <- data.frame(group,subgroup,value)
 
# treemap
#treemap(value,
           # index=names(value),
            #vSize="Freq",
          #  type="index",
        #palette="Set1"
        #    ) 
```

```{r}
pokemon %>%
  select_if(is.numeric) %>%
  {
    data.frame(
      col_name = names(.),
      mean = apply(., 2, mean, na.rm = TRUE),
      std_dev = apply(., 2, sd, na.rm = TRUE)
    )
  } %>%
  mutate(
    mean = round(mean, 2),
    std_dev = round(std_dev, 2)
  ) %>%
  knitr::kable(
    row.names = FALSE,
    col.names = c("Column Name", "Mean", "Standard Deviation")
  )
```



|Variable     | Mean                                     | Standard Deviation                     |
|-------------|------------------------------------------|----------------------------------------|
|capture rate |`r mean(pokemon$capture_rate)`            | `r sd(pokemon$capture_rate)`           |
|attack       |`r mean(pokemon$attack)`                  | `r sd(pokemon$attack)`                 |                           
|defense      |`r mean(pokemon$defense)`                 | `r sd(pokemon$defense)`                |  
|hp           |`r mean(pokemon$hp)`                      | `r sd(pokemon$hp)`                     |  
|height       |`r mean(pokemon$height_m, na.rm = TRUE)`  | `r sd(pokemon$height_m, na.rm = TRUE)` |  
|weight       |`r mean(pokemon$weight_kg, na.rm = TRUE)` | `r sd(pokemon$weight_kg, na.rm = TRUE)`|
|speed        |`r mean(pokemon$speed, na.rm = TRUE)`     | `r sd(pokemon$speed, na.rm = TRUE)`    |  
