---
title: "Untitled"
format: pdf
---

```{r}
library(tidyverse)
```

For the most part, the data is tidy. There are a few columns that are exceptions.

`abilities` is a list of abilities the pokemon has.

`japanese_name` is actually the japanese ways represented in romaji and then katakana.

`capture_rate` has one row with multiple values. This has resulted in the entire column being encoded as a string.

For the Japanese names, we can use the following cell to split the column into two different ones.

```{r}
pokemon <- read_csv("pokemon.csv") %>%
  mutate(
    katakana_name = japanese_name %>% str_extract(r"([^[a-zA-Z0-9?]]+)"),
    romaji_name = japanese_name %>% str_extract(r"([a-zA-Z0-9?]+)"),
    .keep = "unused"
  )
```

It may help to see the case where `capture_rate` can't be encoded into an integer.
̛
```{r}
pokemon %>%
  filter(capture_rate %>% str_detect(r"( )")) %>%
  select(
    contains("name"),
    contains("type"),
    generation,
    is_legendary,
    capture_rate,
    hp
  )
```

Here, this is due to the gen 7 pokemon Minior having two different caputure rates depending on an item that can be equipped.

```{r}
pokemon %>%
  filter(name != "Minior") %>%
  mutate(capture_rate = as.numeric(capture_rate))
```

According to the [site where the data was scraped](https://www.serebii.net/games/capture.shtml), we have the following functions to calculate whether or not the pokemon is captured.

```{r}
catch_value <- function(max_hp, curr_hp, catch_rate, ball_modifier = 1, status_modifier = 1)
   (( 3 * max_hp - 2 * curr_hp ) * (catch_rate * ball_modifier) / (3 * max_hp) ) * status_modifier

catch <- function(x) {
  1048560 / sqrt(sqrt(16711680 / x))
}
```

If `catch_value` returns 255 or higher, then the pokemon is caught. Otherwise, `catch` is applied to the returned value of `catch_value` and then a random number is generated. If the random number is bigger than catch, the pokemon is caught. Now it may help to see what the values of `catch` can be.

```{r}
0:255 %>%
  {ggplot(mapping = aes(., catch(.)))} +
  geom_point() +
  labs(
    x = "Catch Value",
    y = "Catch"
  )

catch(0:255) %>%
  summary()
```

Here, the values seem to be in the range of a 16-bit unsigned integer. Given the output of `catch`, the event where a pokemon is captured follows a bernoulli distribution with probability of success that is the output of `catch` divided by $2^16$.

A probability of success vector function will be made.

```{r}
p_catch <- function(x)
  ifelse(
    x < 255, # check if x < 255 element wise
    (65535 * sqrt(sqrt(x / 255)) * 2^-16), # if so compute probability of success with formula
    1 # otherwise success
  )
```

I now use this to compute the minimum and maximum probabilities of capture for the pokemon and subset the dataframe to be more tidy.


```{r}
prob_df <- pokemon %>%
  select(
    !starts_with(c("against", "type")),
    -abilities
  ) %>%
  filter(name != "Minior") %>%
  mutate(
    capture_rate = as.numeric(capture_rate),
    p_catch_min = catch_value(hp, hp, capture_rate) %>% p_catch(),
    p_catch_max = catch_value(hp, 1, capture_rate) %>% p_catch()
  )

summary(prob_df)

ggplot(prob_df) +
  geom_point(aes(hp, p_catch_min), color = "green", alpha = 0.25) +
  geom_point(aes(hp, p_catch_max), color = "red", alpha = 0.25) +
  labs(
    title = "min = green, max = red",
    x = "Max HP",
    y = "Probability of Capture"
  )
```

It may help to look at the minimum and maximum probabilities of capture.

```{r}
ggplot(prob_df, aes(p_catch_min, p_catch_max)) +
  geom_point() +
  labs(
    x = "Minimum Probability of Capture",
    y = "Maximum Probability of Capture"
  )
```

This seems to indicate that the exteme values for probability of capture have a strong linear relationship.

